---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NeuralNetIndicator

<!-- badges: start -->
<!-- badges: end -->

The goal of NeuralNetIndicator is to provide functionality to calculate a financial trading indicator based
on artificial neural network price predictions. The model bases price forecasts on reference stock prices of the same
sector. The focus is on the utilization of lag and time delays in price movements of similar stocks in order to 
profitably forecast price changes. The information contained in the indicator values and suggestions of how to take
advantage of those calculations can be found under the subheadings *Indicator Values* and 
*Optimization and Backtesting* at the end of this document. 
<br/><br/>
Along the walk-forward trained model with various training-algorithm 
adjustments and tweaking possibilities, this also contains a function to download historical stock prices into a 
suitable xts object as well as functionality to plot the oscillating indicator along the underlying stock prices as a
chart. 

## Installation

In order to install NeuralNetIndicator from GitHub, you will first need to install devtools.

``` r
# if not installed
install.packages("devtools")

devtools::install_github("n1tecki/NeuralNetIndicator")
```

## Example

This is an example which shows you how to access historical data and train the forecast-model:

```{r example}
library(NeuralNetIndicator)
## basic example code
```

### getReferenceSymbols

First, the **getReferenceSymbols** downloads historical data into the xts object *symbols*. 
Along the stock prices of interest, reference assets (preferably from the same (GISC) sector) 
should be listed with their stock symbols. In this case a time span of approximately 20 years will be used.

```{r message = FALSE, warning = FALSE}
symbols <- getReferenceSymbols(x = c('XOM','BP','COP','OXY','PXD','TOT','VLO'), start_date = '2000-10-10', 
                               end_date = '2020-10-10')
head(symbols, 4)
```

### walkForwardNeuralnet

The xts object *symbols* of closing price of all stocks is used as the input for the **walkForwardNeuralnet**. 
The traded asset at focus is *XOM*, hence the first column is selected for the parameter *trad_col*. 
With a in-sample window of 700 days and out-of-sample of 60, the parameter *slid* - number of days of how far into 
the future the forecast should be calculated - is selected to be 4.\
The closing prices of *XOM*, price predictions and indicator value for each day are exported into the xts object 
*indicator*.

```{r message = FALSE}
indicator <- walkForwardNeuralnet(symbols, trad_col=1, neurons_int = c(3,5), 
                                  in_sample = 700, out_sample = 60, slid = 4)
head(indicator, 4)
```

### indicatorChart

The **indicatorChart** function allows to plot a chart of the indicator, along the stock prices of interest 
(*symbol*) and *zoom* in on points of interest. The indicator values *indicator$indicator* from the previously
received *indicator* object are passed on, along the stock code of the corresponding stock prices.

```{r message = FALSE, echo = TRUE}
indicatorChart(indicator$indicator, symbol = 'XOM', zoom = '2018-07-30::2019-07-01')
```

It is also possible just to plot the indicator itself to visualize the values.

```{r message = FALSE, echo = TRUE}
indicatorChart(indicator$indicator, zoom = '2018-07-30::2019-07-01')
```

## Indicator Values
The oscillating values of the indicator are the percentage difference between the closing price of that day and
the price forecast for *slid* days ahead calculated by the neural network.\
For example, an indicator value of 1.4
means that the neural network estimates a 1.4% increase in price in *slid* days. The same goes for negative values -
when the indicator shows a value of -2.3, the forecast is that the price will fall 2.3%.\
<br/><br/>
Based on those values buy or sell positions can be generated. For example if the indicator reaches a value of 2, a long position is entered; when the value is above 3, the long position is increased. 

## Optimization and Backtesting
As the **walkForwardNeuralnet** function has many parameters that can be changed, it is encouraging the user to test
different combinations of parameter values (in-sample/out-of-sample length, number of nodes, different combination 
of stocks, back-propagation algorithm of the neural network, error threshold, ...) for their profitability. One
recommendation is to use a genetic algorithm and back-testing the different combinations of parameters for their
profitability.


